---
title: "Exam01"
format: html
---

**Grand Research Question: **
What does the conusmption of each food category in each country look like? 

```{r}
library(tidyverse)
library(tidytuesdayR)
```

```{r}
tuesdata <- tt_load('2020-02-18')
```


```{r}
dim(tuesdata[["food_consumption"]])

food_consumption <- tuesdata[["food_consumption"]]

head(food_consumption, 10)

class(food_consumption)
```

**Understand Variables Individually**
```{r}
ggplot(food_consumption, aes(y=country)) +
  geom_bar()
```


```{r}
ggplot(food_consumption, aes(y=food_category)) +
  geom_bar()
```

```{r}
ggplot(food_consumption, aes(x=consumption)) +
  geom_density()
```

Count of each country and food category the same across all data points; consumption in general lies around the 15 point. 

**Understand Consumption**

```{r}
ggplot(food_consumption, aes(fill=food_category, x=consumption)) + 
  geom_density(alpha = 0.5)
```

```{r}
#| fig-height: 22
#| fig-width: 11
ggplot(food_consumption, aes(fill=country, x=consumption)) + 
  geom_density(alpha = 0.4)
```
These data are so large that it is difficult to visualize them. 

**Answering Grand RQ**
```{r}
#| fig-height: 22
#| fig-width: 22
ggplot(food_consumption, aes(fill=country, x=consumption)) +
  geom_density(alpha = 0.4) +
  facet_wrap(~ "food_category")
```
I struggled quite a bit with finding an appropriate viz as the data points being examined are so large













