---
title: "Exam 2: Food Consumption 2"
format: html
---

## Background

```{r}
library(tidyverse)
library(tidytuesdayR)
library(rnaturalearth)
library(sf)
```

Get data
```{r}
food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-02-18/food_consumption.csv')
```

```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```

Inspect data
```{r}
str(fc)
```
```{r}
head(fc, 22)
```

Fix food category variable
```{r}
food_consumption <- food_consumption |>
  mutate(fc = food_category) |>
  select(!food_category) |>
  mutate(fc = str_replace(fc, "Lamb & Goat", "Lamb")) |> 
  mutate(fc = str_replace(fc, "Milk - inc. cheese", "Dairy")) |> 
  mutate(fc = str_replace(fc, "Wheat and Wheat Products", "Wheat")) |> 
  mutate(fc = str_replace(fc, "Nuts inc. Peanut Butter", "Nuts")) 
```

## Most Consuming Countries

### Research Question 1: Which 5 countries consume the most food? 

```{r}
food_consumption |>
  group_by(country) |>
  summarize(total_consumption = sum(consumption, na.rm = TRUE)) |>
  arrange(desc(total_consumption)) |>
  head(5)
```
```{r}
food_consumption |>
  group_by(country) |>
  summarize(total_consumption = sum(consumption, na.rm = TRUE)) |>
  arrange(desc(total_consumption)) |>
  head(5) |>
  mutate(country = fct_relevel(country, c("Albania", "Netherlands", "Sweden", "Lithuania", "Finland"))) |>
ggplot(aes(y = country, x = total_consumption, fill = country)) +
  geom_col()
```

## Most Consuming Countries of each food

### Research Question 2: What are the top 3 countries that consume each food category? 

```{r}
food_consumption |>
  arrange(desc(consumption)) |>
  group_by(fc) |>
  slice_max(order_by = consumption, n=5) |>
  select(fc, country, consumption)
```

```{r, fig.width= 10}


food_consumption |>
  arrange(desc(consumption)) |>
  group_by(fc) |>
  slice_max(order_by = consumption, n=5) |>
ggplot(aes(x= consumption, y = country)) +
  geom_col() +
  facet_wrap(~ fc, scales = "free_y")
```

## Food Consumption

### Research question 3: What does the consumption of each food look like? 

See provided chloropleth maps 





