---
title: "Effective Viz"
format: html
---

## Warm Up

Recall: Benefits of Visualizations

Understand what we’re working with:
scales & typical outcomes
outliers, i.e. unusual cases
patterns & relationships
Refine research questions & inform next steps of our analysis.
Communicate our findings and tell a story.

### No one right viz

There is no one right way to visualize a data set, eg, check out the 100 ways used to visualize one dataset: https://100.datavizproject.com/ The visualized data was featured in this TidyTuesday!

**Activity: Plot Critique**
In groups:

Scroll through the plots.
Identify at least 1 plot that you feel illuminates some important aspect of the data.
# 3 illuminates change
Identify at least 1 plot that does NOT illuminate the data in a good way.
#28 is really confusing 

## Ugly, Bad, Wrong Viz

One way to identify effective viz is to understand what makes a viz ineffective. In the Fundamentals of Data Visualization, Wilke breaks down ineffective viz into 3 categories:

**Wrong**

The viz is “objectively incorrect”, as in the numbers / trends being displayed are wrong.

**Bad**

The viz is “unclear, confusing, overly complicated, or deceiving”.

**Ugly**

The viz correct and clear but The aesthetics are problematic.

**Activity: Critical Analysis**
Let’s try some critical analysis on specific examples. For your assigned viz, identify the following:

The story the viz is trying to communicate.
Whether the viz is good, ugly, bad, wrong, or some combination.
Areas for improvement.

## Exercises

### Exercise 1: Professionalism

```{r}
# Load tidyverse package for plotting and wrangling
library(tidyverse)

# Import the data
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))
```

Basic plot: 

```{r}
ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point()
```

#### Part a

Add professional labels: 

```{r}
ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point() + 
  labs(x = "Temperature at 9 AM", y = "Temperature at 3 PM", title = "Daily Temperatures by Location", color = "Location")  
```

#### Part b

When we’re including our plot in an article, paper, book, or other similar outlet, we should (and are expected to) provide a more descriptive figure caption. Typically, this is instead of a title and is more descriptive of what exactly is being plotted.

 - Add a figure caption in the top of the chunk.
 - Include your x-axis, y-axis, and legend labels from Part a.
 - Render your Rmd and check out how the figure caption appears.

```{r fig.cap = "Predictability of 3PM temperature by 9AM temperature in Hobart, Uluru, and Wollongong"}

ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point() + 
  labs(x = "Temperature at 9AM", y = "Temperature at 3PM", color = "Location")  
```


### Exercise 2: Accessibility

```{r}
ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "3pm temperature (Celsius)")  
```

#### Part a

Alt text should concisely articulate:

 - What your visualization is (e.g. a density plot of 3pm temperatures in Hobart, Uluru, and Wollongong, Australia).
 - A 1-sentence description of the most important takeaway.
 - A link to your data source if it’s not already in the caption.



```{r}
#| fig-cap: "Density plots of 3pm temperatures in 3 Australian locations."
#| fig-alt: "Density plot of 3pm temperatures in Hobart, Uluru, and Wollongong, Australia. Figure shows that the average temperature in Hobart is Approximately 15 degrees celcius, average in Uluru is approximately 35 degrees celcius, and the average in Wollangong is approximately 20 degrees celcius. Data pulled from https://mac-stat.github.io/data/weather_3_locations.csv"

ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "3pm temperature (Celsius)")  
``` 
 
#### Part b

Color accessibility: 

Using the website "https://www.color-blindness.com/coblis-color-blindness-simulator/" we can examine how if color is being used as a factor in a visualization, some viewers my be unable to decipher between colors and thus would not be able to understand the viz. 
#### Part c

```{r}
ggplot(weather, aes(x = temp3pm, fill = location)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "3pm temperature (Celsius)") + 
  scale_fill_viridis_d()    
```

```{r}
# In the color scale line:
# Change "fill" to "color" since we use color in the aes()
# Change "d" (discrete) to "c" (continuous) since maxtemp is on a continuous scale
ggplot(weather, aes(y = temp3pm, x = temp9am, color = maxtemp)) + 
  geom_point(alpha = 0.5) + 
  labs(x = "3pm temperature (Celsius)") + 
  scale_color_viridis_c()
```

### Exercise 3: Ethics

Consideration of *impact*

#### Part a

**Climate change plot:** this plot is unethical because the scale used on the y axis is too large to properly visualize the data. This plot would lead viewers to think there was no change in the anverage annual global temperature since 1880. 

#### Part b

**We need to consider: visibility, privacy, power, emotion & embodiment, pluralism, & context**

Pick one or more of the following examples of data viz to discuss with your group. How do the approaches taken:

 - emphasize one or more of: visibility, privacy, power, emotion, embodiment, pluralism, and/or context?
 - improve upon what we might be able to convey with a simpler bar chart, scatterplot, etc?

#### Part c
Read: https://data-feminism.mitpress.mit.edu/

### Exercise 4: Critique

Practice critiquing some more complicated data viz listed at Modern Data Science with R, Exercise 2.5.

Think about the following questions:

What story does the data graphic tell? What is the main message that you take away from it?
Can the data graphic be described in terms of the Grammar of Graphics (frame, glyphs, aesthetics, facet, scale, guide)? If so, please describe.
Critique and/or praise the visualization choices made by the designer. Do they work? Are they misleading? Thought-provoking? Are there things that you would have done differently?


### Exercise 5: Design Details

Go in deeper, refer to class website
























