---
title: "Bivariate Viz"
---

Use this file for practice with the **bivariate viz** in-class activity.  Refer to the class website for details.


---
title: "Bivariate Viz"
---

Use this file for practice with the **bivariate viz** in-class activity.  Refer to the class website for details.

```{r}
# Import data
survey <- read.csv("https://hash-mac.github.io/stat112site-s25/data/survey.csv")

library(tidyverse)

# How many students have now filled out the survey?


# What type of variables do we have?
```

```{r}
# Attach a package needed to use the ggplot function
ggplot(survey, aes(x=fav_temp_c)) +
  geom_histogram(color="lightgreen", fill="pink", binwidth=10) +
  theme_minimal() 

# Make a ggplot
```

## ICA Exercises

```{r}
# Load data
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

# Check it out
head(elections)
```

# Exercise 0

```{r}
ggplot(elections, aes(x=winner_20)) +
  geom_bar()
```

```{r}
ggplot(elections, aes(x=repub_pct_20)) +
  geom_histogram()
```

## Exercise 1
```{r}
#ggplot(elections, aes(x=repub_pct_16)) +
  #geom_dotplot(y=repub_pct_20)
```

## Exercise 2

```{r}
# Set up the plotting frame
# How does this differ than the frame for our histogram of repub_pct_20 alone?
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_text(aes(label = state_abbr), color="pink")

```

## Exercise 3

strong, positive relationship; TX has somewhat higher occurrence in the lower percentages as compared to general trend

## Exercise 4

```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth()
```


Without original points: 
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_smooth()
```


Part b: "best" *linear model*

```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Exercise 5

```{r}
ggplot(elections, aes(y=median_rent, x=repub_pct_20)) +
  geom_smooth(method="lm")
```
  
```{r}
ggplot(elections, aes(y=median_age, x=repub_pct_20)) +
  geom_smooth(method="lm")
```
  
## Exercise 6: Sad Scatterplot

```{r}
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_point()
```
  
There are better ways to visualize/ compare categorical and quantitative data

## Exercise 7: Quantitative vs Categorical – Violins & Boxes

```{r}
# Side-by-side violin plots
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_violin()
```
  
```{r}
# Side-by-side boxplots (defined below)
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_boxplot()
```
So: we can visualize that it is less common for blue --> more common for red, with outliers, etc

## Exercise 8: Quantitative vs Categorical – Intuition Check

```{r}
ggplot(elections, aes(x = repub_pct_20)) +
  geom_density()

ggplot(elections, aes(x = repub_pct_16)) +
  geom_density()

# idk
```

## Exercise 9: Quantitative vs Categorical – Density Plots

```{r}
# Name two "bad" things about this plot
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density()
```

Colors overlap, colors don't match the listed "colors" of the states

```{r}
# What does scale_fill_manual do?
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

```{r}
# What does alpha = 0.5 do?
# Play around with different values of alpha, between 0 and 1
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

```{r}
# What does facet_wrap do?!
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red")) +
  facet_wrap(~ historical)
```

```{r}
# Let's try a similar grouping strategy with a histogram instead of density plot.
# Why is this terrible?
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_histogram(color = "white") +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

## Exercise 10

Which do you like best?

- box or density

What is one pro of density plots relative to boxplots?

- larger, clearer, overlappable

What is one con of density plots relative to boxplots?

- no median, mean, or specific outliers

## Exercise 11: Categorical vs Categorical – Intuition Check

```{r}
# Plot 1: adjust this to recreate the top plot
ggplot(elections, aes(x = historical)) +
  geom_bar()
```

## Exercise 12: Categorical vs Categorical

```{r}
# A stacked bar plot
# How are the "historical" and "winner_20" variables mapped to the plot, i.e. what roles do they play?
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
```

```{r}
# A faceted bar plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar() +
  facet_wrap(~ historical)
```

```{r}
# A side-by-side bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "dodge")
```

```{r}
# A proportional bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "fill")
```

## Exercise 13: Practice

```{r}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv")
```

```{r}
# How do 3pm temperatures (temp3pm) differ by location?

ggplot(weather, aes(x = temp3pm, fill = location)) +
  geom_density(alpha = 0.5) 
```

```{r}
ggplot(weather, aes(y = temp3pm, x = temp9am)) + 
  geom_smooth() 
```

```{r}
ggplot(weather, aes(x = location, fill=raintoday)) +
  geom_bar()
```



