---
title: "Mulivariate Viz"
---

Use this file for practice with the **mulivariate viz** in-class activity.  Refer to the class website for details.

```{r}
library(tidyverse)

weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") |> 
  mutate(date = as.Date(date))  
```


Check out the first 6 rows
# What are the units of observation?


# How many data points do we have? 
nrow(weather) --> 2367
# What type of variables do we have?
str(weather --> data frame; num, chr, int, 

```{r}
head(weather)
```

Example 1
```{r}
ggplot(weather, aes(x=temp3pm, y=location)) +
  geom_point()
```
I got this wrong, i shouldve contructed a desity plot of only temp3pm

## Example 2
```{r}
# Plot 1 (no facets & starting from a density plot of temp3pm)
ggplot(weather, aes(x = temp3pm)) + 
  geom_density()
```
## Example 2
```{r}
ggplot(weather, aes(x = temp3pm, fill=location)) + 
  geom_density(alpha=0.5)
```
```{r}
#plot 2
# Plot 2 (no facets or densities)
ggplot(weather, aes(y = temp3pm, x = location)) + 
  geom_boxplot()
```

```{r}
# Plot 3 (facets)

ggplot(weather, aes(x = temp3pm, fill=location)) + 
  geom_bar()

```
I also got this one wrong; I should've constructed 3 side-by-side density graphs

```{r}
# Don't worry about the syntax (we'll learn it soon)
woll <- weather |>
  filter(location == "Wollongong") |> 
  mutate(date = as.Date(date))  
```

## Example 3 
```{r}
# How often does it raintoday?
# Fill your geometric layer with the color blue.
ggplot(woll, aes(x = raintoday)) +
  geom_bar(fill="lightblue") 
  
```

What does it tell us about rain tomorrow? 
```{r}
ggplot(woll, aes(x = raintoday, y=raintomorrow)) +
  geom_smooth() 
```
I did this wrong; I should've constructed bar graphs 


```{r}
# Now compare different approaches

# Default: stacked bars
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar()
```

```{r}
# Side-by-side bars
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "dodge")
```

```{r}
# Proportional bars
# position = "fill" refers to filling the frame, nothing to do with the color-related fill
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "fill")
```

```{r}
# Instead of a curve that captures the general TREND,
# draw a line that illustrates the movement of RAW temperatures from day to day
# NOTE: We haven't learned this geom yet! Guess.
ggplot(woll, aes(y = temp3pm, x = date))
  geom_smooth()
```
I don't know what to do here

answer: 
ggplot(woll, aes(y = temp3pm, x = date)) + 
  geom_point() + 
  geom_smooth(span = 0.5)
  
And for raw: 
```{r}
ggplot(woll, aes(y = temp3pm, x = date)) + 
  geom_line()
```
## Exercise 5
```{r}
# Plot temp3pm vs temp9am
# Change the code in order to indicate the location to which each data point corresponds
ggplot(weather, aes(y = temp3pm, x = temp9am,color=location)) + 
  geom_point()
```
I have no idea 
Answer: 
```{r}
ggplot(weather, aes(y = temp3pm, x = temp9am, color = location)) + 
  geom_point()
```

\

Example 6
```{r}
# Change the code in order to construct a line plot of temp3pm vs date for each separate location (no points!)
ggplot(weather, aes(y = temp3pm, x = date)) +
  geom_line()
```

I still have no idea
Answer:
ggplot(weather, aes(y = temp3pm, x = date, color = location)) + 
  geom_line()

Exercise 7
```{r}
# Plot the relationship of raintomorrow & raintoday
# Change the code in order to indicate this relationship by location
ggplot(weather, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "fill")
```
nothing

Answer: 
ggplot(weather, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "fill") + 
  facet_wrap(~ location)

### Exercises

```{r}
# Import and check out data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")
head(education)
```

Part a
```{r}
ggplot(education, aes(x=sat)) +
  geom_density()
```

Part b
Range 800 to 1100
Typical: 900 and to a lesser extent 1050
Shape: two humps (I forgot the name)

## Exercise 2
Part a
```{r}
ggplot(education, aes(x=sat, y=expend)) +
  geom_smooth(method= "lm")
```
```{r}
ggplot(education, aes(x=sat, y=salary)) +
  geom_smooth(method= "lm")
```
Part b

Interesting that these are both plots with negative relatiobships, it seems like more fundint would mean better scores, but its the opposite, and the same for teacher salary.

## Exercise 3
```{r}
ggplot(education, aes(x=sat, y=expend, color=salary)) +
  geom_point()
```

## Exercise 4

```{r}
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 2))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```

```{r}
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 3))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```

## Exercise 5
### Part a

```{r}
ggplot(education, aes(x=fracCat)) + 
  geom_bar()
```

### Part b
```{r}
ggplot(education, aes(x=sat, y=fracCat)) +
  geom_point()
```
```{r}
ggplot(education, aes(x=sat, color=fracCat)) +
  geom_density()
```



### Part c
```{r}
ggplot(education, aes(x=sat, y=expend, color=fracCat)) + 
  geom_point() +
  geom_smooth(method="lm")
```

The stats with less expenditure also took fewer SATs, so the people who do take it stand out as outstanding students; **simpson's paradox**









